Texture mapping 3D planar models of indoor environments with noisy camera poses

### 2013-03-03 spie/paper.tex
first revision of revisions

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{abstract}
  Automated 3D modeling of building interiors is used in applications
  such as virtual reality and environment mapping. Texturing these
  models allows for photo-realistic visualizations of the data
  collected by such modeling systems. While data acquisition times for
  mobile mapping systems are considerably shorter than for static
  ones, their recovered camera poses often suffer from inaccuracies,
  resulting in visible discontinuities when successive images are
  projected onto a surface for texturing. Existing methods to stitch
  images together are often computationally expensive and work
  independently of pose estimates and geometry data. We present a
  method for texture mapping that reduces complexity by selecting
  images whose camera poses can be align in 2D. We perform alignment
  of images to geometry as well as images to each other, which
  produces visually consistent textures even in the presence of
  inaccurate surface geometry. We also present two different methods
  for selecting and compositing images into a single texture,
  depending on the geometry of the surface being textured. The
  effectiveness of these methods is demonstrated on a number of different
  indoor environments.
\end{abstract}

@@

The first step in automated 3D modeling is the physical scanning of
the environment's geometry. An indoor modeling system must be able to
recover camera poses within an environment while simultaneously
reconstructing the 3D structure of the environment itself
\cite{chen2010indoor, hz, kua2012loopclosure, liu2010indoor}. This is
known as the simultaneous localization and mapping (SLAM) problem, and
is generally solved by taking readings from laser range scanners,
cameras, and inertial measurement units (IMUs) at multiple locations
within the environment.

Mounting such devices on a platform carried by an ambulatory human
provides unique advantages over vehicular-based systems on wheels in
terms of agility and portability, but can also result in larger
localization error \cite{liu2010indoor}. As a result, common methods
for texture mapping generally produce poor results.

In this paper, we present a number of approaches to texture mapping 3D
models of indoor environments made of planar surfaces in the presence
of uncertainty and noise in camera poses. In particular, we consider
data obtained from a human-operated backpack system with a number of
laser range scanners as well as 2 cameras facing left and right, each
equipped with fisheye lenses reaching an approximately 180$^{\circ}$
field of view and taking photos at a rate of 5 Hz. Applying multiple
localization and loop-closure algorithms on the raw data collected by
the onboard sensors \cite{chen2010indoor, kua2012loopclosure,
  liu2010indoor}, the backpack is localized\footnote{In this paper, we
  use the terms localization and pose recovery interchangeably, in
  that they both refer to recovering position and orientation.}  over
its data collection period. This requires recovering the 6 degrees of
freedom for the backpack as well as the cameras rigidly mounted on
it. Once this is complete, the data from the laser range scanners is
used to generate a 3D point cloud of the surrounding environment, from
which a 3D planar model is created \cite{sanchez2012point}. This
model, consisting of 2D polygonal planes in 3D space, along with the
set of images captured by the backpack's cameras and their noisy 3D
poses, can be considered the input to our texture mapping problem.

We perform texture mapping on each planar surface independently and in
parallel. For each surface, we begin by obtaining a set of images that
spans the entire surface with high resolution imagery. We then use our
noisy camera poses to project these selected images onto the
surface. These projections are then refined in 2D, in order to
maximally align them with the surface's geometry, as well as to each
other, allowing us to handle both errors in geometry as well as camera
poses. For surfaces and camera poses at any location or orientation,
we demonstrate a simple tile-based approach for sampling
high-resolution portions of images and compositing them into a
texture. In cases where cameras have consistently perpendicular
viewing angles, we demonstrate a superior approach that leads to more
seamless textures.

The remainder of the paper is organized as follows. Section
\ref{sec:simpleTextureMapping} explains how images are projected onto
geometry and shows a simple texture mapping procedure. Section
\ref{sec:existingApproaches} covers existing approaches to image
stitching, and their performance on our datasets. Section
\ref{sec:2dAlignment} contains our approach towards efficient 2D image
alignment, followed by Section \ref{sec:imageCompositing}, which
describes two methods of selecting and compositing images to create
the final texture. Section \ref{sec:resultsAndConclusions} contains
results and conclusions.

@@

\section{Simple Texture Mapping}
\label{sec:simpleTextureMapping}

Before describing our texture mapping procedure, we use this section
to explain basic texture mapping and a method for selecting images for
texturing. The geometry of the texture mapping process for a planar
surface is shown in Figure \ref{fig:projection}.  Given a set of $M$
images to texture a target plane, camera matrix $P_i$ for the $ith$
image, translates a 3D point in the world coordinate system to a 2D
point or pixel in image $i$'s coordinates. A camera matrix $P_i$ is
composed of the camera's intrinsic parameters, containing focal length
and image center, as well as extrinsic parameters which specify the
rotation and translation of the camera's position in 3D world
coordinates at the time that image $i$ is taken. These extrinsic
parameters are determined by the backpack hardware and localization
algorithms \cite{chen2010indoor, liu2010indoor, kua2012loopclosure}
and are substantially noisy.

Because the backpack system takes pictures at a rate of 5 Hz, hundreds
of images are available for texturing each surface in the model. Our
objective in designing a texture mapping process is to determine which
of these images should be used, and where their contents should map
onto the final texture, in order to minimize any visual
discontinuities or seams that would suggest that the plane's texture
is not composed of a single continuous image. In the remainder of this
section, we provide an overview of a simple tile-based texture mapping
procedure.

@@

Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate,
the most straightforward approach is to discretize the target plane
into small square tiles, and choose an image to texture each tile
directly. We choose to work with rectangular units to ensure that
borders between any two distinct images in the final texture are
either horizontal or vertical. Since most strong environmental
features inside buildings are horizontal or vertical, any visible
seams in the texture intersect them minimally and are less noticeable.

In order to select an image for texturing a tile $t$, we first gather
a list of candidate images that contain all four of its corners, which
we can rapidly check by projecting $t$ into each image using the $P_i$
camera matrices. Furthermore, each candidate image must have been
taken at a time when its camera had a clear line-of-sight to $t$,
which can be determined using standard ray-polygon intersection tests
between the camera location, $t$, and every other surface
\cite{rayintersection}.

Once we have a list of candidate images for $t$, we define a scoring
function in order to objectively select the best image for texturing
$t$. Since resolution decreases and camera pose errors become more
pronounced with distance, we wish to minimize the distance between
cameras and the surfaces they texture. Additionally, we desire images
that are projected perpendicularly, rather than obliquely, onto the
plane, maximizing the resolution and amount of useful texture
available in their projections, as well as minimizing any parallax
effects due to real-world geometry not accurately represented by the
digital model. In other words, we wish to minimize the angle between
the tile's normal vector and the camera axis for images selected for
texturing that tile. These two criteria can be met by maximizing the
function $\frac{1}{d} (-1 \cdot \vec{c}) \cdot \vec{n}$ as shown in
Figure \ref{fig:scoringFunction}, where $d$ is the distance between
the centers of a camera and a tile, and $\vec{n}$ and $\vec{c}$ are
the directions of the plane's normal and the camera axis respectively.

As Figure \ref{fig:compareAll}(a) demonstrates, this approach results
in image boundaries with abrupt discontinuities between tiles, due to
significant misalignment between images, though most of it appears to
be reconcilable using 2D transforms. Given the large number of input
images, it is possible to develop a more sophisticated image selection
procedure, as described in Section \ref{sec:imageCompositing},
however, a more significant and reliable improvement can be found
through first improving the alignment of the images as shown in the
next 2 sections.

\begin{figure}[H]
  \centering \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_naive.jpg}}
  \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_naive_shift.jpg}}

@@

  \caption{(a): Tile-based texturing. (b): Tile-based texturing after
    image alignment. (c): Tile-based texturing after image alignment
    with caching. (d): Shortest path texturing after image
    alignment). (e,f): Blending applied to (c) and (d). (g,h): Zoomed
    in views of discontinuities in (e) vs. in (f).}
  \label{fig:compareAll}
\end{figure}

\section{Existing Approaches to Image Alignment}
\label{sec:existingApproaches}
Stitching together multiple images to produce a larger, seamless image
is a commonly performed task, with many successful approaches over the
past years. Generally, parts of images are matched to each other,
usually through direct comparisons or feature detection and
matching. Images are then transformed to maximize matches, often by
calculating homographies between pairs of images, or by iteratively
adjusting camera poses in 1 to 6 degrees of freedom.

Feature matching has a number of advantages over direct matching that
make it more suitable for our often non-planar data, and rotational
differences \cite{szeliski2006image}. Feature matching however, works
best when multiple unique visual references exist in the environment
that can be detected in multiple images. In contrast, indoor
environments have a high prevalence of bare surfaces, as well as
repeating textures, such as similar windows, doors, and wall-mounted
decorations, that cause difficulty in disambiguating features. This
lack of reliable reference points often results in errors when
matching images together.

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=6in]{graphApproach.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=6in]{autostitchResult.jpg}}

  \centering \subfloat[][]{\includegraphics[width=6in]{finalLong.jpg}}

In prior work, we experimented with integrating image stitching with
an iterative localization algorithm \cite{liu2010indoor}. When run on
long chains of images, especially where features are sparse, this
approach produces distorted textures, as seen in Figure
\ref{fig:mosaic3D}(a). Furthermore, this approach is not closed-form,
and its iterative camera adjustment process over large datasets led to
prohibitively long computation time.

The AutoStitch software package performs homography-based alignment as
well, with additional provisions that attempt to reduce drift and
increase efficiency \cite{panorama2d, autostitch}. Though AutoStitch
also performs well in areas with dense features, it can not handle
areas without features, and has trouble aligning wall sections with
even short segments of bare texture. The example in Figure
\ref{fig:mosaic3D}(b) was generated after many rounds of manual
tuning, and areas with fewer visual features or repeating texture
patterns simply failed outright with AutoStitch.

\section{2D Image Alignment}
\label{sec:2dAlignment}
In this section, we describe our method for efficient and robust image
alignment. Rather than attempt to register all of our images in 3D, as
many state-of-the-art techniques for image stitching do, we instead
select a subset of images that can be aligned in 2D. The subset of
images we select for alignment and subsequent texturing corresponds to
the images selected by the tile-based texturing procedure described in
Section \ref{sec:simpleTextureMapping}.

We have found that applying 2D alignments to this set of images works
well for a few reasons. First, the nature of our input data and the
selected images is such that localization error chiefly occurs in two
dimensions, which correspond to the plane of the surface being
projected onto. This is because our backpack operator, during data
acquisition, makes efforts to walk within a few meters of, and
parallel to all walls being scanned. As a result, the translational
error of camera poses is quite minimal in the dimension perpendicular
to the surface being textured. Furthermore, because our set of images
was selected to have minimal distance from our surface, rotational
errors have a lesser effect, except for rotation around the axis
perpendicular to the surface being textured. Therefore, the majority
of both our translation and rotational errors occur in the 2
dimensions of the surface's plane. This of course does not apply for
ceilings and floors; however, the lack of features on most ceilings
and floors means that 3D misalignment does not often have a visible
effect, beyond what can be fixed in 2D.

Our proposed 2D alignment procedure consists of three parts. First, images
are projected onto the surface and lines within these projections are
detected. These lines are then used to transform images such that they match geometry-based lines composing the boundaries of the surface being textured. Second, occlusion checks are performed to remove invalid parts of each image for the target surface. Third, SIFT feature matches are detected between pairs of images, and a weighted linear least squares problem is solved in order to maximize all image and geometry-based alignments. Each step will now be explained in detail.

@@

\begin{figure}
  \centering
  \caption{}
  \label{fig:geometryAlignment}
\end{figure}

After computing each image's projection onto the target surface, as
described in Section \ref{sec:simpleTextureMapping}, we obtain a set
of image-based line segments by using Hough transforms to detect lines
in the image projections. We then gather a set of geometry-based
lines, which correspond to the lines comprising the target surface's
border, as well as lines formed where other surfaces intersect the
target surface. An example of these lines is shown for a ceiling
surface in Figure \ref{fig:geometryAlignment}. Given perfect camera
poses and surface geometry, the lines in images corresponding to
corners between surfaces should match up exactly with borders and
intersections of surfaces. By inducing such lines to match, we fit our
camera poses more accurately to our surface, and therefore to
eachother.

To perform geometry-based alignment, we collect pairs of image-based
and geometry-based lines, which are within a distance and orientation
difference threshold of each other. We have found a distance threshold
of 250 mm and an orientation difference threshold of $10^\circ$ to
work well for our datasets. In practice, we rarely have more than 2
such pairs, but if so, we select the 2 noncollinear pairs with the
most similar orientation difference and continue. With 2 noncollinear
pairs, we select the pair with the longer image-based line, and
perform a rotation of the image such that the lines in that pair
become parallel. We then perform a translation such that the lines in
that same pair overlap. This translation has ambiguity in the
dimension along that line, which is resolved by matching the midpoint
of the image-based line in the second pair to its corresponding
geometry-based line. Thus, with 2 or more pairs, we can obtain a fixed
rotation and translation, which we save for usage in Section
\ref{sec:robustSIFTFeatureMatching}. If we only have 1 pair, we
perform a rotation and a minimal translation to match its lines. This
translation's ambiguity however, can not be resolved, but is saved
also to be used in Section
\ref{sec:robustSIFTFeatureMatching}. Finally, in the case where we
have no matches, we can still perform rotations in order to exploit
patterns in indoor environments. For instance, doors, windows,
furniture, etc. all tend to have linear edges that are parallel to the
edges of the surfaces they are on. Similarly, lights, visible interior
walls, etc. which are visible in floor and ceiling images, all tend to
be parallel to exterior walls corresponding to the edges of ceiling
and floor surfaces. Thus, we choose to minimize the angle between
image-based lines and any geometry-based line. We do this using the
RANSAC framework, which successfully ignores outliers
\cite{fischler1981random}.

After these steps are performed, our image projections line up well with our target surface, as shown in Figure \ref{fig:geometryAlignment}. This procedure reconciles both errors in camera poses as well as in geometry, and results in sharp, continuous borders across images, which is crucial when performing occlusion checks.

\subsection{Image Occlusion}
\label{sec:imageOcclusion}
In order to properly texture surfaces such as the one shown in Figure
\ref{fig:occlusion}, it is important to detect when parts of image
projections contain texture belonging to other surfaces and should
therefore not be used. This can be tested by recursively performing
ray-polygon intersection tests between camera locations and other
surfaces in a regularly spaced grid \cite{rayintersection}. Where four
corners of a rectangular region are occluded, texture is
removed. Where no corners are occluded, the recursion stops. Where
there is a mixture of both, the rectangular region is subdivided into
four, and the same process is performed on each. By performing Section
\ref{sec:geometryAlignment}'s alignment procedure before occlusion,
texture belonging to other surfaces is accurately removed, which is
necessary for the next section.

\subsection{2D Feature Alignment}
\label{sec:robustSIFTFeatureMatching}
Our next step is to align images to each other by searching for
corresponding feature points between all pairs of overlapping
images. We use feature alignment rather than pixel or intensity-based
alignment due to the differences in lighting as well as possible
occlusion among our images, both of which feature alignment is less
sensitive to \cite{lowe1999object, mikolajczyk2005performance,
  szeliski2006image}.  We use SiftGPU \cite{siftgpu} for its high
performance on both feature detection as well as pairwise
matching. These matches determine $dx$ and $dy$ distances between each
pair of features for two image projections, though these distances may
not always be the same for different features. Since indoor
environments often contain repetitive features such as floor tiles or
doors, we need to ensure that SIFT-based distances are
reliable. First, we only perform alignment on parts of images that
overlap given the original noisy poses. Second, we discard feature
matches that correspond to an image distance greater than 200 mm from
what the noisy poses estimate. In order to utilize the remaining
feature matches robustly, RANSAC \cite{fischler1981random} is again
used to estimate the optimal $dx_{i,j}$ and $dy_{i,j}$ distances
between two images $i$ and $j$. We use a 5 mm threshold for RANSAC, so
that SIFT matches are labeled as outliers if their distance is not
within 5 mm of the sampled average distance.

We now use the feature-based distances between each pair of images as
well as our previous geometry alignment results to refine all image
positions using a weighted linear least squares solution. The
variables we wish to solve for are the $x_i$ and $y_i$ positions of
images, while equations are the feature-based distances between pairs
of images, images fixed to geometry with 0 or 1 degrees of freedom,
and the original noisy camera poses. The original camera poses are
needed because we sometimes are not able to detect feature matches in
all of our images, or we lack enough geometry alignment results to
generate a single solution. Because we want to minimally use our
original poses, we give them a weighting factor of $0.01$, while all
other equations are weighted at $1$. An example setup for a weighted
linear least squares problem $\textrm{min}_{\vec{\beta}}
||W^\frac{1}{2}(A \vec{\beta} - \vec{\gamma})||_2^2 $ with 3 images is
shown below.

\[
A =
\begin{pmatrix}
  -1 & 1 & 0 & 0 & 0 & 0\\
  0 & 0 & 0 & -1 & 1 & 0\\
  0 & -1 & 1 & 0 & 0 & 0\\
  0 & 0 & 0 & 0 & -1 & 1\\
  0 & -m_2 & 0 & 0 & 1 & 0\\
  1 & 0 & 0 & 0 & 0 & 0\\
  0 & 0 & 0 & 1 & 0 & 0\\
  1 & 0 & 0 & 0 & 0 & 0\\
  0 & 0 & 0 & 1 & 0 & 0\\

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{abstract}
  Automated 3D modeling of building interiors is used in applications
  such as virtual reality and environment mapping. Texturing these
  models allows for photo-realistic visualizations of the data
  collected by such modeling systems. While data acquisition times for
  mobile mapping systems are considerably shorter than for static
  ones, their recovered camera poses often suffer from inaccuracies,
  resulting in visible discontinuities when successive images are
  projected onto a surface for texturing. We present a method for
  texture mapping that starts by selecting images whose camera poses
  are well-aligned in two dimensions. We then align images to geometry
  as well as to each other, producing visually consistent textures
  even in the presence of inaccurate surface geometry and noisy camera
  poses. The final step is to select and composite images into a
  texture mosaic. The effectiveness of the proposed method is
  demonstrated on a number of different indoor environments.
\end{abstract}

@@

The first step in automated 3D modeling is the physical scanning of
the environment's geometry. An indoor modeling system must be able to
recover its pose within an environment while simultaneously
reconstructing the 3D structure of the environment itself
\cite{chen2010indoor, hz, kua2012loopclosure, liu2010indoor}. This is
known as the simultaneous localization and mapping (SLAM) problem, and
is generally solved by taking readings from laser range scanners,
cameras, and inertial measurement units (IMUs) at multiple locations
within the environment. Mounting such devices on a platform carried by
an ambulatory human provides unique advantages over vehicular-based
systems on wheels in terms of agility and portability, but can also
result in larger localization error \cite{liu2010indoor}. As a result,
common methods for texture mapping generally produce poor results.

In this paper, we present an approach to texture mapping 3D models of
indoor environments made of planar surfaces in the presence of
uncertainty and noise in camera poses. In particular, we consider data
obtained from a human-operated backpack system with a number of laser
range scanners as well as 2 cameras facing left and right, each
equipped with fisheye lenses reaching an approximately 180$^{\circ}$
field of view and taking photos at a rate of 5 Hz. Applying multiple
localization and loop-closure algorithms on the raw data collected by
the onboard sensors \cite{chen2010indoor, kua2012loopclosure,
  liu2010indoor}, the backpack is localized\footnote{In this paper, we
  use the terms localization and pose recovery interchangeably, in
  that they both refer to recovering position and orientation.}  over
its data collection period. This requires recovering the 6 degrees of
freedom for the backpack itself as well as the cameras rigidly mounted
on it. Once this is complete, the data from the laser range scanners
is used to generate a 3D point cloud of the surrounding environment,
from which a 3D planar model is created \cite{sanchez2012point}. This
model, consisting of 2D polygonal planes in 3D space, along with the
set of images captured by the backpack's cameras and their noisy 3D
poses, can be considered the input to our texture mapping problem.

\begin{figure}
  \centering
  \includegraphics[trim=0cm 3cm 0cm 0cm, clip=true,
  width=6in]{flowchart.jpg}
  \caption{The proposed texture mapping procedure\\}
  \label{fig:flowchart}
\end{figure}

The overall block diagram for the proposed texture mapping procedure
is shown in Figure \ref{fig:flowchart}. We texture map each planar
surface independently and in parallel. For each surface, we begin by
obtaining a set of images that spans the entire surface with high
resolution imagery. We then use our noisy camera poses to project
these selected images onto the surface. These projections are then
refined in 2D, in order to maximally align them with the surface's
geometry, as well as to each other, allowing us to handle both errors
in geometry as well as camera poses. For surfaces and camera poses at
arbitrary locations or orientations, generally ceilings and floors, we
propose a tile-based approach for sampling high-resolution portions of
images and compositing them into a texture. In cases where cameras
have consistently perpendicular viewing angles to the surfaces under
consideration, generally walls, we demonstrate a superior approach
that leads to more seamless textures.

The remainder of the paper is organized as follows. Section
\ref{sec:existingApproaches} covers existing approaches to image
stitching, and their performance on our datasets. Section
\ref{sec:simpleTextureMapping} explains how to downsample the set of
available images by selecting those with the best orientation and
distance from surfaces under consideration. Section
\ref{sec:2dAlignment} contains our approach towards 2D image
alignment, followed by Section \ref{sec:imageCompositing}, which
describes two methods of selecting and compositing images to create
the final texture. Sections \ref{sec:results} and \ref{sec:conclusion}
contain results and conclusions.

@@

\section{Existing Approaches to Image Alignment}
\label{sec:existingApproaches}
There are many existing approaches to stitching together multiple
images to produce a larger, seamless image \cite{szeliski2006image,
  agarwalapanoramas, wangmultipleviews, coorg1997matching,
  debevechybrid, bernardinimultiplescans}. Generally, parts of images
are matched to each other, either through direct comparisons or
feature detection and matching. Images are then transformed to
maximize matches, often by computing homographies between pairs of
images, or by iteratively adjusting camera poses in 1 to 6 degrees of
freedom.

Feature matching has a number of advantages over direct matching that
make it more suitable for our often non-planar environments, and
rotational differences between camera poses
\cite{szeliski2006image}. Feature matching however, works best when
multiple unique visual references exist in the environment that can be
detected in multiple images. In contrast, indoor environments have a
high prevalence of bare surfaces, as well as repeating textures, such
as similar windows, doors, and wall-mounted decorations, that are
difficulte to disambiguate. This lack of reliable reference points
often results in errors when matching images together.

\begin{figure}
  \centering \subfloat[][]{\includegraphics[trim=0cm 0cm 20cm 0cm,
    clip=true, width=6in, height=0.7in]{graphApproach.jpg}}

  \centering \subfloat[][]{\includegraphics[trim=0cm 0cm 15cm 0cm,
    clip=true, width=6in, height=0.7in]{autostitchResult.jpg}}

  \centering \subfloat[][]{\includegraphics[trim=0cm 0cm 75cm 0cm,
    clip=true, width=6in, height=0.7in]{finalLong.jpg}}

In prior work, we have experimented with integrating image stitching
with the iterative global localization algorithm used to recover the 6
degrees of freedom of the backpack acquisition system
\cite{liu2010indoor}. When run on long chains of images, especially
where features are sparse, this approach produces distorted textures,
as seen in Figure \ref{fig:mosaic3D}(a). Furthermore, this approach is
not closed-form, and its iterative camera adjustment process over
large datasets led to prohibitively long computation time.

The AutoStitch software package performs homography-based alignment as
well, with additional provisions that attempt to reduce drift and
increase efficiency \cite{panorama2d, autostitch}. Though AutoStitch
performs well in areas with dense features, it can not handle areas
without features, and has trouble aligning wall sections with even
short segments of bare texture. The example in Figure
\ref{fig:mosaic3D}(b) has been generated after many rounds of manual
tuning; we have empirically found that areas with fewer visual
features or repeating texture patterns simply failed outright with
AutoStitch.

\section{Image Selection}
\label{sec:simpleTextureMapping}

The geometry of the texture mapping process for a planar surface is
shown in Figure \ref{fig:projection}. Given a set of $M$ images to
texture a target plane, camera matrix $P_i$ for the $ith$ image
transforms a 3D point in the world coordinate system to a 2D point or
pixel in image $i$'s coordinates. A camera matrix $P_i$ is composed of
the camera's intrinsic parameters, containing focal length and image
center, as well as extrinsic parameters which specify the rotation and
translation of the camera's position in 3D world coordinates at the
time that image $i$ is taken. These extrinsic parameters are
determined by the backpack hardware and the corresponding localization
algorithms \cite{chen2010indoor, liu2010indoor, kua2012loopclosure}
and are noisy.

Since the backpack system takes pictures at a rate of 5 Hz, hundreds
of images are available for texturing each surface in the 3D
model. Our objective in designing a texture mapping process is to
determine which of these images should be used, and where their
contents should map onto the final texture, in order to minimize any
visual discontinuities or seams that would suggest that the plane's
texture is not composed of a single continuous image. In the remainder
of this section, we propose an image selection criteria and use it to
derive a simple tile-based texture mapping procedure.

@@

Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate,
a straightforward approach for texturing is to discretize the target
plane into small square tiles, and choose an image to texture each
tile directly. We choose to work with rectangular units to ensure that
borders between any two distinct images in the final texture are
either horizontal or vertical. Since most environmental features
inside buildings are horizontal or vertical, any visible seams in the
texture intersect them minimally and are therefore less noticeable.

In order to select an image for texturing a tile $t$, we first gather
a list of candidate images that contain all four of its corners; we
can rapidly check this by projecting $t$ into each image using the
$P_i$ camera matrices. Furthermore, each candidate image must have
been taken at a time when its camera had a clear line-of-sight to $t$,
which can be determined using standard ray-polygon intersection tests
between the camera location, $t$, and every other surface
\cite{rayintersection}.

Once we have a list of candidate images for $t$, we define a scoring
function in order to objectively select the best image for texturing
$t$. Since resolution decreases and camera pose errors become more
pronounced with distance, we wish to minimize the distance between
cameras and the surfaces they texture. Additionally, we desire images
that are projected perpendicularly, rather than obliquely, onto the
plane, maximizing the resolution and amount of useful texture
available in their projections, as well as minimizing any parallax
effects due to real-world geometry not accurately represented by the
digital 3D model. In other words, we wish to minimize the angle
between the tile's normal vector and the camera axis for images
selected for texturing that tile. These two criteria can be met by
maximizing the function $\frac{1}{d} (-1 \cdot \vec{c}) \cdot \vec{n}$
as shown in Figure \ref{fig:scoringFunction}, where $d$ is the
distance between the centers of a camera and a tile, and $\vec{n}$ and
$\vec{c}$ are the directions of the plane's normal and the camera axis
respectively. As Figure \ref{fig:compareAll}(a) demonstrates, this
approach, when used directly for texturing, results in image
boundaries with abrupt discontinuities between tiles, due to
misalignment between images, though most of it appears to be
reconcilable using 2D transforms.

While this procedure can be used to generate a texture, it is also
useful for obtaining a collection of images that can be used as a
starting point for our proposed texturing procedure, as explained in
Section \ref{sec:2dAlignment}. For our datasets, this process
generally selects around 10\% of all the images that could be used for
texturing the surface. This roughly 10:1 subsampling not only reduces
computational complexity of the remaining steps, but also selects the
most preomising images for the remaining steps.

Given the large number of input images, it is possible to develop a
more sophisticated image selection procedure, as described in Section
\ref{sec:imageCompositing}, however, a more significant and reliable
improvement can be found through first improving the alignment of the
images as shown in the next section.

\begin{figure}
  \centering \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_naive.jpg}}
  \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_naive_shift.jpg}}

@@

  \caption{(a) Tile-based texturing; (b) Tile-based texturing after
    image alignment; (c) Tile-based texturing after image alignment
    with caching; (d) Shortest path texturing after image
    alignment; (e,f) Blending applied to (c) and (d); (g,h) Zoomed
    in views of discontinuities in (e) vs. in (f).}
  \label{fig:compareAll}
\end{figure}

\section{2D Image Alignment}
\label{sec:2dAlignment}
In this section, we describe our method for efficient and robust image
alignment. Rather than register all of our images in 3D, as many
state-of-the-art techniques for image stitching do, we instead select
a subset of images to be aligned in 2D. This subset corresponds to the
images selected by the image selection procedure described in Section
\ref{sec:simpleTextureMapping}.

Applying 2D alignments to this set of images works well for a number
of reasons. First, the nature of our input data and the selected
images is such that localization error chiefly occurs in two
dimensions, which correspond to the plane of the surface being
projected onto. This is because the backpack operator, during data
acquisition, makes efforts to walk within a few meters of, and
parallel to all walls being scanned. As a result, the translational
error of camera poses is quite minimal in the dimension perpendicular
to the wall surface being textured. Furthermore, because our set of
images has minimal distance from its corresponding wall surface,
rotational errors are less pronounced, except for rotation around the
axis perpendicular to the surface being textured. Therefore, most
translation and rotational errors occur in the 2 dimensions of the
wall surface's plane. This of course does not apply for ceilings and
floors; however, the lack of features on most ceilings and floors
means that 3D misalignment is not likely to visually make a
difference, beyond what can be done in 2D.

Our proposed 2D alignment procedure consists of three parts. First,
images are projected onto the surface and lines within these projected
images are detected. Images are then transformed such that these lines
match geometric lines composing the boundaries of the surface being
textured. Second, occlusion checks are performed to remove invalid
parts of each image for the target surface. Third, SIFT feature
matches are detected between pairs of images, and a weighted linear
least squares problem is solved in order to maximize all image and
geometry-based alignments. Each step will now be explained in detail.

@@

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=6in]{allunshifted.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=6in]{allshifted.jpg}}

  \centering \subfloat[][]{\includegraphics[trim=0cm 3cm 0cm 0cm,
    clip=true, width=5in]{matchlines.jpg}}

  \caption{Images projected onto a ceiling surface, where
    geometry-based lines corresponding to the ceiling's boundary are
    shown in red. Image-based lines detected by Hough transform in the
    image projections are shown in blue; (a) images projected with
    their original noisy camera poses; (b) after image alignment to
    maximize line matches between images and geometry; (c) examples of
    matching lines in cases with $\geq$ 2 line pairs, 1 line pair, and
    zero line pairs, from top to bottom.}
  \label{fig:geometryAlignment}
\end{figure}

After computing each image's projection onto the target surface, as
described in Section \ref{sec:simpleTextureMapping}, we obtain a set
of image-based line segments by using Hough transforms to detect lines
in the image projections. We then gather a set of geometry-based
lines, which correspond to the lines comprising the target surface's
border, as well as lines formed where other surfaces intersect the
target surface. An example of these lines is shown in red for a
ceiling surface in Figure \ref{fig:geometryAlignment}(a). Ideally, for
perfect camera poses and surface geometry, the lines in images
corresponding to corners between surfaces should match up exactly with
corners in the 3D model. By inducing such lines to match, we fit
camera poses more accurately to the surface, and to each other as
well.

To align images to surface geometry, we collect pairs of image-based
and geometry-based line segments, which are within a distance and
angular difference threshold of each other. We have found a distance
threshold of 250 mm and an orientation difference threshold of
$10^\circ$ to work well for our datasets. For each pair of lines, we
compute the angular difference between the pair's image and geometry
lines. If there are 2 or more pairs with angular differences within
$1^\circ$, we select the two pairs with the longest noncollinear
image-based lines, and rotate the image such that the lines in the
pair with the longest image-based line become parallel. We then find a
translation such that the lines in that same pair overlap. This
translation has ambiguity in the dimension along the matched lines,
which is resolved by matching the midpoint of the image-based line in
the second pair to its corresponding geometry-based line. Thus, with 2
or more pairs, it is possible to obtain a fixed rotation and
translation, which are saved for usage in Section
\ref{sec:robustSIFTFeatureMatching}.

If there are not 2 or more pairs with similar angular differences, we
select the pair with the longest image-based line and apply a rotation
and the minimal translation to match its lines. This translation's
ambiguity however, can not be resolved, but is also saved to be used
in Section \ref{sec:robustSIFTFeatureMatching}. Finally, in the case
where there are no line pairs, we can still rotate images in order to
exploit patterns in indoor environments. For instance, doors, windows,
furniture, etc. tend to have linear edges that are parallel to the
edges of the surfaces they are on. Similarly, lights, visible interior
walls, etc. which are visible in floor and ceiling images, tend to be
parallel to exterior walls corresponding to the edges of ceiling and
floor surfaces. Thus, we choose to minimize the angle between
image-based lines and geometry-based lines. We use the RANSAC
framework, with an inlier difference of $0.5^\circ$ in order to
successfully ignore outliers \cite{fischler1981random}.

After these steps, image projections line up well with
target surfaces, as shown in Figure \ref{fig:geometryAlignment}(b),
which is considerably more aligned than Figure
\ref{fig:geometryAlignment}(a) is. This procedure reconciles both
errors in camera poses as well as in geometry, and results in sharp,
continuous borders across images, which is crucial when checking for
occlusion.

\subsection{Image Occlusion}
\label{sec:imageOcclusion}

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=3in]{occlusiondiagram.png}}
  \hspace{0.4cm} \subfloat[][]{\includegraphics[width=3in,
    height=1.8in]{occlusionimages.png}}
  \caption{(a) The image from the camera in this diagram contains
    texture that belongs to the grey occluding surface, which should
    not be projected onto the orange target surface; (b) without
    geometry alignment, texture to the left of the red line would be
    removed, which would leave some erroneous texture projected onto
    our target surface; (c) after geometry alignment, the image is
    shifted, resulting in the correct amount of texture being
    removed.}
  \label{fig:occlusion}
\end{figure}

In order to correctly texture surfaces, it is important to detect and
remove parts of image projections containing texture for occluding
surfaces. For instance, in Figure \ref{fig:occlusion}(a), an image
used to texture the orange target surface also contains part of a grey
occluding surface. We remove this incorrect texture by recursively
performing ray-polygon intersection tests between the camera location
and other surfaces \cite{rayintersection}. These intersection tests
are performed at the corners of a grid overlaid upon the target
surface. Where all four corners of a grid section are occluded,
texture is removed. Where one or more corners are occluded, the grid
is subdivided into four, and the process repeats. Occlusion checking
works entirely with geometry, so by ensuring that images match
geometry using \ref{sec:geometryAlignment}'s alignment procedure,
texture belonging to other surfaces is accurately removed, as seen in
Figure \ref{fig:occlusion}(b).
